---
title: "pracma_fminsearch"
author: "Kirk Gosik"
date: "December 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(pracma)
library(lme4) # for Orange dataset
rm(list=ls()); gc(reset = TRUE)

logi_fit <- function(params, data){
  a <- params[1]
  b <- params[2]
  r <- params[3]
  
  t <- data$age # time variable
  y <- data$circumference
  
  y_hat <- a/(1+b*exp(-r*t))
  sum((y - y_hat)^2)
}
  
out <- fminsearch(logi_fit, c(100, 1, 1), data = Orange)

logi_predict <- function(out, data){
  a <- out$xval[1]
  b <- out$xval[2]
  r <- out$xval[3]
  
  t <- data$age
  
  y_hat <- a/(1+b*exp(-r*t))
  y_hat
}

```

Adding a fixed effect for a group variable

```{r}
Orange_group <- data.frame(Orange, group = c(rep(1, 20), rep(0, 15)))

logi_fit_group <- function(params, data){
  a <- params[1]
  b <- params[2]
  r <- params[3]
  beta <- params[4]
  
  t <- data$age # time variable
  x <- data$group
  y <- data$circumference
  
  y_hat <- a/(1+b*exp(-r*t)) + beta*x
  sum((y - y_hat)^2)
}
  
out <- fminsearch(logi_fit_group, c(100, 1, 1, 1), data = Orange_group)
out

```

# Adding legendre polynomicals without correlated error structure

 with the currenct output by including a group (or SNP) effect will result in the need for a paramter for the order of the legendre polynomials + 1

For finding the order of the legendre polynomial, you can use AIC or BIC in some fashion. 

```{r}
Orange_group <- data.frame(Orange, group = c(rep(1, 20), rep(0, 15)))

logi_fit_group <- function(params, data){
  a <- params[1]
  b <- params[2]
  r <- params[3]
  beta <- params[4:7]
  
  t <- data$age # time variable
  x <- data$group
  y <- data$circumference
  
  t_sd <- (t - min(t))/(max(t) - min(t))
  L <- legendre(3, t_sd)
  
  y_hat <- a/(1+b*exp(-r*t)) + beta %*% (L*x)
  sum((y - y_hat)^2)
}
  
out <- fminsearch(logi_fit_group, c(100, 1, 1, 1, 1, 1, 1), data = Orange_group)
out
```

Correlating the error structure

```{r}
Orange_group <- data.frame(Orange, group = c(rep(1, 20), rep(0, 15)))

logi_fit_cor <- function(params, data){

  a <- params[1]
  b <- params[2]
  r <- params[3]
  sigma <- params[4] # for use in ar1 model
  rho <- params[5] # for use in ar1 model
  
  id <- data$Tree
  j <- length(unique(id))
  t <- data$age # time variable
  m <- length(unique(t))
  #x <- data$group
  y <- data$circumference
 
  COVAR <- rho^(outer(seq(0, (m-1)), seq(0, (m-1)), function(a,b) abs(a-b)))
  diag(COVAR) <- sigma
  COVAR <- kronecker(diag(j), COVAR)
  
  y_hat <- a/(1 + b * exp(-r * t))
  
  #sum((y - y_hat)^2)
  t(y - y_hat) %*% COVAR %*% (y - y_hat)
}
  
out <- fminsearch(logi_fit_cor, c(150, 700, 350, 0.1, 0.1), data = Orange, dfree = FALSE)

# might need constraints to properly optimze
## look into optimx package/function

```



## PRACMA

```{r}
library(pracma)
library(MASS)
library(magrittr)

# Sigma <- matrix(c(1, 0.2, 0.04, 
#                   0.2, 1, 0.2,
#                   0.04, 0.2, 1), 3, 3)
# mvrnorm(10, mu = c(1, 1, 1), Sigma = Sigma)

rand_logi_func <- function( a = 1, 
                            b = 1, 
                            r = 1, 
                            t, 
                            sigma = 0.5, 
                            rho = 0.2, 
                            type = "ar1" ){
  
  mu_vec <- a /( 1 + b * exp(-r * t) )

  SIGMA = matrix(0, nrow = length(mu_vec), ncol = length(mu_vec))
  diag(SIGMA) <- sigma^2
  
  rho <- switch(type,
                `cs` = rep(rho, (length(mu_vec) - 1)),
                `ar1` = sapply(1 : length(mu_vec), function(i) rho^(i-1))
  )
  
  for( i in 1 : nrow(SIGMA) ){
    for( j in 1 : ncol(SIGMA) ){
     if( i != j ) SIGMA[i, j] <- rho[i] 
    }
  }
  
  mvrnorm(1, mu = mu_vec, Sigma = SIGMA)
 
}

rand_logi_func(a = 115,
               b = 2.5,
               r = .5,
               t = 1:15,
               sigma = 1,
               rho = 0.2,
               type = "ar1") %>%
  plot

```

# Legendre

```{r}
# from Wolfram Mathworld

P0 <- function(x){ 1 }
P1 <- function(x){ x }
P2 <- function(x){ 0.5 * (3*x^2 - 1) }
P3 <- function(x){ 0.5 * (5*x^3) - 3*x }
P4 <- function(x){ 0.125 * (35*x^4 - 30 * x^2 + 3) }
P5 <- function(x){ 0.125 * (63*x^5 - 70 * x^3 + 15 * x) }
P6 <- function(x){ 0.0625 * (231*x^6 - 315 * x^4 + 105 * x^2 - 5) }

#print(legendre)

t <- seq(0, 15, len = 10)
t_s <- t/15

(L <- legendre(3, t_s))
(y_t <- as.matrix(rand_logi_func(a = 115,
               b = 2.5,
               r = .5,
               t = t,
               sigma = 1,
               rho = 0.2,
               type = "ar1")))
plot(y_t)

L %*% y_t

```
```{r}

t <- Orange_df$age
t_s <- (t - min(t))/(max(t)- min(t)) # min-max standardization
y_t <- Orange_df$circumference
L <- legendre(3, t_s)
L %*% y_t # might not need to do this part, that is what we are fitting


library(tidyr)
Orange_df <- data.frame(Orange)
t <- unique(Orange_df$age)
t_s <- (t - min(t))/(max(t)- min(t))# min-max standardization
y_t <- as.matrix(spread(Orange_df, Tree, circumference)[,-1])
L <- legendre(3, t_s)
out <- L %*% y_t   # might not need to do this part, that is what we are fitting
out
rowSums(out) # same as above
rowMeans(out) # average across subjects
gather(as.data.frame(out)) # subject specific

```

## fminsearch Linear Regression

```{r fminsearch, echo=FALSE}
# Rosenbrock function
rosena <- function(x, a) 100*(x[2]-x[1]^2)^2 + (a-x[1])^2 # min: (a, a^2)
fminsearch(rosena, c(-1.2, 1), a = sqrt(2))

# Sample data
n <- 100
x1 <- rnorm(n)
x2 <- rnorm(n)
y <- 1 + x1 + x2 + rnorm(n)
X <- cbind( rep(1,n), x1, x2 )
dat <- data.frame(x1,x2,y)
summary(fit_lm <- lm(y ~ ., data = dat))


leastsquares <- function(data, beta ){
  x <- as.matrix(cbind(1, data[,-3]))
  y <- data[,3]
  
  y_pred <- x %*% beta
  sum((y - y_pred)^2)
  
}


min.RSS <- function(data, par) {
              with(data, sum((par[1] + par[2] * x1 + par[3] * x2 - y)^2))
}

fit_lm
fminsearch(leastsquares, c(0,1,1), data = dat)
fminsearch(min.RSS, c(0,1,1), data = dat)


```

## fminsearch Non-Linear Regression

```{r}

n <- 100
x <- rnorm(n)
y <- 1 * (1 - exp(-0.5 * x)) + rnorm(n)
dat <- data.frame(x,y)

# y = a0(1-exp(-a1x))

non_lin <- function(beta, data){
  x <- data$x
  y <- data$y
  
  y_pred <- beta[1] * (1 - exp(-beta[2] * x))
  sum((y - y_pred)^2)
}

fminsearch(non_lin, c(1,1), data = dat)



```



```{r}
y_test <- rand_logi_func(a = 115,
               b = 2.5,
               r = .5,
               t = 1:15,
               sigma = 1,
               rho = 0.2,
               type = "ar1")

dat <- data.frame(y = y_test,
                  t = 1:15)

ar1_nonlin <- function(params, data){
  y <- data$y
  t <- data$t
  
  a <- params[1]; b <- params[2]; r <- params[3]
  sigma <- params[4]; rho <- params[5]
  
  y_hat <- 
  
  
}

```

